---
title: "Vic LRE manuscript"
author: "Jia Qi Beh"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Victoria LRE manuscript - analysis

This document contains all R packages and codes used in the bioinformatic analyses of the manuscript 'Genomics to understand the global landscape of linezolid resistance in Enterococcus faecium and Enterococcus faecalis'.


# Genome download

Download Illumina reads from NCBI SRA using kingfisher (https://github.com/wwood/kingfisher-download).

```{bash}
# Create ENA accessions file.
# Run kingfisher
for line in $(cat accessions.txt); do kingfisher get -r ${line} -m ena-ascp aws-http prefetch --download-threads 10 -f fastq.gz -t 8; done
```


# fastp
version 0.23.2

```{bash}
# create accessions.txt
# run fastp
for line in $(cat accessions.txt); do fastp --thread 16 -r -W 4 -M 20 -l 36 -i ${line}_1.trim.fastq.gz -I ${line}_2.trim.fastq.gz -o ${line}_R1_trimmed.fastq.gz -O ${line}_R2_trimmed.fastq.gz -j ${line}.json -h ${line}.html ; done
```


# short-read assembly
spades version 3.15.5 (using shovill)

```{bash}
# create accession.txt
# run shovill
while read line; do nice shovill --outdir ${line} --R1 ${line}_R1_trimmed.fastq.gz --R2 ${line}_R2_trimmed.fastq.gz --cpus 16; done < accessions.txt
```


# long-read assembly
porechop version 0.2.4
filtlong version 0.2.1
canu version 2.2
unicycler version 0.5.0
dragonflye version 1.1.0

Run porechop
```{bash}
porechop -i input_reads.fastq.gz -o output_reads.fastq.gz
```

Run filtlong
```{bash}
filtlong --min_length 1000 --keep_percent 90 --target_bases 500000000 input.fastq.gz | gzip > output.fastq.gz
```

Run canu
```{bash}
canu -d ID -p ID genomeSize=3m -trimmed -nanopore ID_output.fastq.gz maxThreads=16
```

Run unicycler
```{bash}
unicycler -1 ID_R1.trimmed.fastq.gz -2 ID_R2.trimmed.fastq.gz -l ID_output.fastq.gz -o OUTPUT -t 12
```

Run dragonflye
```{bash}
dragonflye --gsize 3M --cpus 12 --outdir OUTPUT --reads ID.fastq.gz
```


# quast
version 5.2.0

```{bash}
# create accession.txt
# run quast
for line in $(cat accessions.txt); do ${line}/contigs.fa -o ${line} -t 16; done
```


# prokka
version 1.14.5

```{bash}
# create accession.txt
# run prokka
for line in $(cat accessions.txt); do prokka --outdir ${line} --prefix ${line} --cpus 16 ${line}/contigs.fa; done
```


# kraken2
version 2.1.3

```{bash}
# make accessions file
# run kraken2 with default database
kraken2 --output ${line} --report output.txt assembly.fa
```


# mlst
version 2.23.0

```{bash}
mlst --scheme efaecium assembly.fa --threads 10
```


# abritamr
version 1.0.11

```{bash}
abritamr run --contigs input.txt --species Enterococcus_faecium -j 16
abritamr run --contigs input.txt --species Enterococcus_faecalis -j 16
```


# lrefinder
version 1.0

```{bash}
python3 LRE-Finder.py -ipe ID_R1_trimmed.fastq.gz ID_R2_trimmed.fastq.gz -o ID -t_db lre-finder/elmDB/elm -ID 80 -1t1 -cge -matrix
```


# mob-typer
mobsuite version 3.0.3

```{bash}
mob_typer --infile assembly.fa --out_file output.txt
```


# bakta
version 1.8.1

```{bash}
bakta --db path/to/bakta/db assembly.fa
```


# snippy + iqtree
version 4.4.5

```{bash}
# run snippy
snippy --cpus 16 --outdir output --ref reference.gb --R1 ID_R1.trimmed.fastq.gz --R2 ID_R2.trimmed.fastq.gz

# run snippy-core
snippy-core AUSMDU* --ref reference.gbff

# run iqtree v2.0.6
iqtree -s core.aln -m GTR+I+G -B 1000 -nt 16
```


# rhierbaps
version 1.1.4

```{r}
library(rhierbaps)
library(tidyverse)
library(ggtree)
library(phytools)
library(ape)

snp.matrix <- load_fasta("core.aln.fasta")

hb.results.2 <- hierBAPS(snp.matrix, max.depth = 2, n.pops = 20, n.extra.rounds = Inf, quiet = TRUE)

# Load tree file
iqtree <- ape::read.tree("input.treefile")
iqtree

# Check tree tips and rhierbaps 
iqtree$tip.label
hb.results.2$partition.df$Isolate

hb.results.2$partition.df$Isolate %in% iqtree$tip.label
iqtree$tip.label %in% hb.results.2$partition.df$Isolate

all(iqtree$tip.label %in% hb.results.2$partition.df$Isolate)
iqtree$tip.label[!iqtree$tip.label %in% hb.results.2$partition.df$Isolate] # this will print the missing names

# Plot tree
tree.3 <- ggtree(iqtree)
tree.3 <- tree.3 %<+% hb.results.2$partition.df
tree.3 <- tree.3 + geom_tippoint(aes(color = factor(`level 1`)))
tree.3

ggsave("tree.pdf", tree.3, width=8, height=6)
```

