---
title: "OneHealth_visualisation"
author: "Jia Qi Beh"
date: "2026-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Victoria LRE manuscript - visualisation

This document contains all R packages and codes used to create the figures in the manuscript 'Prospective sampling reveals sharing of antimicrobial resistance genes in Enterococcus across One Health compartments'.


# R Libraries

```{r}
pacman::p_load(
  tidyverse,     # data management and visualization
  rio,
  UpSetR,        # special package for combination plots
  ggupset,
  ComplexUpset,
  readxl,
  readr,
  reshape2,      # melt() function
  svglite,
  hrbrthemes,
  viridis,
  RColorBrewer,
  pheatmap,
  cowplot,
  patchwork,
  micropan,
  matrixStats,
  ggforce, 
  reshape2,
  ggtree,
  here,
  ape,
  treeio,
  ggnewscale,
  phytools,
  phangorn,
  pals,
  scales,
  lubridate,
  igraph,
  networkD3,
  htmlwidgets,
  webshot,
  rhierbaps,
  rworldmap,
  rwmGetColours,
  plasmapR)
```

Set working directory
```{r}
setwd("~/Vic-LRE")
```


# Figure 1

## 1a - barplots

E. faecium plot
```{r}
efm <- read.csv("figure_1/efaecium.csv")
efm

# Replace all blank values with 'NA'
efm[efm == ""] <- NA

# Convert percentage
efm.1 <- 
  efm %>%
  count(LzdMIC, LRgenotype, LREvsLSE) %>%
  mutate(pct = n / sum(n) * 100) %>%
  print()

my_colours_efm <- c('#e07a5f', '#457b9d', '#ccc5b9')

plot.efm <- 
  ggplot(efm.1, aes(x = LzdMIC, y = pct, fill=LRgenotype))+ 
  geom_col()+
  scale_fill_manual(values = my_colours_efm)+
  xlab("Linezolid MIC (ug/mL)")+
  ylab("% isolates")+
  theme_bw()+
  theme(strip.text = element_text(face = "italic", size = 10), # Species font
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8, face = "bold"),
        axis.title.x = element_text(face = "bold", size = 8),
        axis.title.y = element_text(face = "bold", size = 8))+
  labs(y=NULL)+  #remove y-axis label
  guides(fill=guide_legend(title="Linezolid genotype"))

plot.efm
```

E. faecalis plot
```{r}
efs <- read.csv("figure_1/efaecalis-only.csv")
efs

# Replace all blank values with 'NA'
efs[efs == ""] <- NA

# Convert percentage
efs.1 <- 
  efs %>%
  count(LzdMIC, LRgenotype, LREvsLSE) %>%
  mutate(pct = n / sum(n) * 100) %>%
  print()

my_colours_efs <- c('#e07a5f', '#ccc5b9')

plot.efs <- 
  ggplot(efs.1, aes(x = LzdMIC, y = pct, fill=LRgenotype))+ 
  geom_col()+
  scale_fill_manual(values = my_colours_efs)+
  xlab("Linezolid MIC (ug/mL)")+
  ylab("% isolates")+
  theme_bw()+
  theme(strip.text = element_text(face = "italic", size = 10), # Species font
        legend.position = "none",
        axis.title.x = element_text(face = "bold", size = 8),
        axis.title.y = element_text(face = "bold", size = 8))+
  guides(fill=guide_legend(title="Linezolid genotype"))

plot.efs
```

Join plots
```{r}
combined_plot <- plot.efs + plot.efm  # arrange side-by-side
combined_plot

ggsave("figure_1.pdf", combined_plot, width=6, height=4, dpi=300)
```

## 1b - upset plot

E. faecalis
```{r}
# Import data file 
data <- read.csv("/Users/jbbeh/Downloads/_VicLRE/figure_1/efs_ggupset.csv")
data

# Reformat LR gene columns from wide to long 
df <- reshape2::melt(data, id.vars = "ID", measure.vars = c("cfr","cfrB","cfrD","optrA","poxtA","G2576T"))  # change from wide to long format
df

df.1 <- df %>% subset(value == "Yes")
df.1 <- df.1[,-3]
df.1 <- df.1 %>% rename(Gene = variable)   # rename column 
df.1

# Group data by Gene
df.2 <- 
  df.1 %>% 
  group_by(ID) %>%
  summarize(Gene = list(Gene))
df.2

# Plot graph 
plot <- 
  df.2 %>%
  ggplot(aes(x = Gene)) +
  geom_bar() +
  scale_x_upset() +
  scale_y_continuous(breaks = c(2,4,6,8,10))   # set y-axis breaks
plot

# Save plot 
ggsave("figure_1b.pdf", plot, width = 8, height = 8)
```

E. faecium
```{r}
# Import data file 
data <- read.csv("/Users/jbbeh/Downloads/_VicLRE/figure_1/efm_ggupset.csv")
data

# Reformat LR gene columns from wide to long 
df <- reshape2::melt(data, id.vars = "ID", measure.vars = c("cfr","cfrB","cfrD","optrA","poxtA","G2576T"))  # change from wide to long format
df

df.1 <- df %>% subset(value == "Yes")
df.1 <- df.1[,-3]
df.1 <- df.1 %>% rename(Gene = variable)   # rename column 
df.1

# Group data by Gene
df.2 <- 
  df.1 %>% 
  group_by(ID) %>%
  summarize(Gene = list(Gene))
df.2

# Plot graph 
plot <- 
  df.2 %>%
  ggplot(aes(x = Gene)) +
  geom_bar() +
  scale_x_upset() +
  scale_y_continuous(breaks = c(2,4,6,8,10))   # set y-axis breaks
plot

# Save plot 
ggsave("figure_1b.pdf", plot, width = 8, height = 8)
```


# Figure 2

Figure 2 was created using BRIG software: https://github.com/happykhan/BRIG


# Figure 3

## 3a - tree

Run rhierbaps
```{r}
snp.matrix <- load_fasta("figure_3/core.aln.fasta")

hb.results.2 <- hierBAPS(snp.matrix, max.depth = 2, n.pops = 20, n.extra.rounds = Inf, quiet = TRUE)
baps_cluster_df <- hb.results.2$partition.df
# write.csv(baps_cluster_df, "efs_baps_cluster.csv", row.names = FALSE)
```

Load tree file
```{r}
iqtree <- ape::read.tree("figure_3/core.aln.fasta")
iqtree

# Check tree tips and rhierbaps 
iqtree$tip.label
hb.results.2$partition.df$Isolate

hb.results.2$partition.df$Isolate %in% iqtree$tip.label
iqtree$tip.label %in% hb.results.2$partition.df$Isolate

all(iqtree$tip.label %in% hb.results.2$partition.df$Isolate)
iqtree$tip.label[!iqtree$tip.label %in% hb.results.2$partition.df$Isolate]

# Plot tree
tree.3 <- ggtree(iqtree)
tree.3 <- tree.3 %<+% hb.results.2$partition.df
tree.3 <- tree.3 + geom_tippoint(size=0.5, aes(color = factor(`level 1`)))
tree.3

# Load metadata
sample_data <- read.csv("/Users/jbbeh/Downloads/_VicLRE/figure_3/ggtree_metadata.csv")
View(sample_data)

# Check whether the Isolate IDs in both files align correctly
sample_data$ID %in% iqtree$tip.label
iqtree$tip.label %in% sample_data$ID 

# Add root and scale bar
tree_mid <- midpoint(iqtree)
tree_mid
```

Make tree with rhierbaps levels
```{r}
tree.3 <- ggtree(tree_mid, branch.length = 0.5) %<+% hb.results.2$partition.df +
  geom_tippoint(size=0.5, aes(color = factor(`level 1`)))
tree.3

# Plot tree with metadata
tree.4 <- tree.3 %<+% sample_data +
  geom_rootedge(rootedge = 0.0005) +   # add root 
  geom_treescale(y=20)
tree.4

# Heatmap 1: MDU
mdu <- data.frame("MDU" = sample_data[,c("MDU")])
rownames(mdu) <- sample_data$ID

h1 <- 
  gheatmap(tree.4, mdu, width = 0.04, color=NULL, offset=0.02,
           colnames = FALSE) +
  scale_fill_manual(name = "continent",
                    values = c("#fb8500","#023047"),
                    na.value="#023047",  # set colour of NA values
                    breaks = c("Yes","No"),
                    labels = c("Yes", "No")) +
  theme(legend.position = "NULL")
h1

# Heatmap 2: continent
length(unique(sample_data$continent))
unique(sample_data$continent) #count total number of continents

continent <- data.frame("continent" = sample_data[,c("continent")])
rownames(continent) <- sample_data$ID

h2 <- h1 + new_scale_fill()
h3 <- 
  gheatmap(h2, continent, width = 0.04, color=NULL, offset=0.03,
           colnames = FALSE) +
  scale_fill_manual(name = "continent",
                    values = c("#f2cc8f", "#f2cc8f","#81b29a","#3d405b","#e07a5f","#fb6f92", "grey"),
                    na.value="grey",  # set colour of NA values
                    breaks = c("asia", "Asia", "Oceania", "South America", "North America", "Europe", ""),
                    labels = c("asia", "Asia", "Oceania", "South America", "North America", "Europe", "")) +
  theme(legend.position = "right")
h3

# Heatmap 3: sample type
length(unique(sample_data$sample_type))
unique(sample_data$sample_type)

sample_data$sample_type <- as.factor(sample_data$sample_type) 
sample_type <- data.frame("sample_type" = sample_data[,c("sample_type")])
rownames(sample_type) <- sample_data$ID

h4 <- h3 + new_scale_fill()
h5 <- 
  gheatmap(h4, sample_type, width = 0.04, color="white", offset=0.04,
           colnames = FALSE) +
  scale_fill_manual(name = "sample type",
                    values = c("#840032","#002642", "#E59500", "#840032", "#840032", "white", "white"),
                    na.value="white",  # set colour of NA values
                    breaks = c("clinical", "animal", "environment","human", "human/clinical", "others"),
                    labels = c("clinical", "animal", "environment","human", "human/clinical", "others"))
h5

# Heatmap 4: ST
## Find top 10 prevalent STs
result <- sample_data %>%
  count(ST) %>%
  arrange(desc(n))
result

sample_data$ST <- as.factor(sample_data$ST) 
sample_data$ST[!(sample_data$ST %in% c("480", "506", "16","476", "631", "21", "368", "314", "585", "256"))] <- 'others'
ST <- data.frame("ST" = sample_data[,c("ST")])
rownames(ST) <- sample_data$ID

h6 <- h5 + new_scale_fill()
h7 <- 
  gheatmap(h6, ST, width = 0.04, color="white", offset=0.05,
           colnames = FALSE) +
  scale_fill_manual(name = "sample type",
                    values = c("#840032","#002642", "#E59500", "#669bbc", "#588157",
                               "#ff99c8", "#83c5be", "#9f86c0", "#4361ee", "#f72585"),
                    na.value="white",  # set colour of NA values
                    breaks = c("480", "506", "16","476", "631", 
                               "21", "368", "314", "585", "256"),
                    labels = c("480", "506", "16","476", "631", 
                               "21", "368", "314", "585", "256"))
h7

# Heatmap 4: LR gene cfr
sample_data$cfr <- as.factor(sample_data$cfr) 
cfr <- data.frame("cfr" = sample_data[,c("cfr")])
rownames(cfr) <- sample_data$ID

h8 <- h7 + new_scale_fill()
h9 <- 
  gheatmap(h8, cfr, width = 0.04, color="white", offset=0.06,
           colnames = FALSE) +
  scale_fill_manual(name = "cfr",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h9

# Heatmap 5: LR gene cfr(B)
sample_data$cfrB <- as.factor(sample_data$cfrB) 
cfrB <- data.frame("cfrB" = sample_data[,c("cfrB")])
rownames(cfrB) <- sample_data$ID

h10 <- h9 + new_scale_fill()
h11 <- 
  gheatmap(h10, cfrB, width = 0.04, color="white", offset=0.07,
           colnames = FALSE) +
  scale_fill_manual(name = "cfrB",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h11

# Heatmap 6: LR gene cfr(D)
sample_data$cfrD <- as.factor(sample_data$cfrD) 
cfrD <- data.frame("cfrD" = sample_data[,c("cfrD")])
rownames(cfrD) <- sample_data$ID

h12 <- h11 + new_scale_fill()
h13 <- 
  gheatmap(h12, cfrD, width = 0.04, color="white", offset=0.08,
           colnames = FALSE) +
  scale_fill_manual(name = "cfrD",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h13

# Heatmap 7: LR gene optrA
sample_data$optrA <- as.factor(sample_data$optrA) 
optrA <- data.frame("optrA" = sample_data[,c("optrA")])
rownames(optrA) <- sample_data$ID

h14 <- h13 + new_scale_fill()
h15 <- 
  gheatmap(h14, optrA, width = 0.04, color="white", offset=0.09,
           colnames = FALSE) +
  scale_fill_manual(name = "optrA",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h15

# Heatmap 8: LR gene poxtA
sample_data$poxtA <- as.factor(sample_data$poxtA) 
poxtA <- data.frame("poxtA" = sample_data[,c("poxtA")])
rownames(poxtA) <- sample_data$ID

h16 <- h15 + new_scale_fill()
h17 <- 
  gheatmap(h16, poxtA, width = 0.04, color="white", offset=0.10,
           colnames = FALSE) +
  scale_fill_manual(name = "poxtA",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))+
  theme(legend.position = "NULL")
h17

ggsave("~/Downloads/plot.pdf", h17, width=8, height=6, dpi = 320)
```


## 3b - barplots

Load data
```{r}
data <- read.csv("figure_3/efm_metadata.csv")
data
```

Count prevalent countries
```{r}
count_country <- 
  data %>% dplyr::count(country) %>%
  arrange(desc(n))
count_country

# Then select only the top 10 countries and change others into "Others"
data.1 <-
  data %>% 
  mutate(newCountry = case_when(
                           country == "Denmark" ~ "Others",
                           country == "Ireland" ~ "Others",
                           country == "South Korea" ~ "Others",
                           country == "Portugal" ~ "Others",
                           country == "Switzerland" ~ "Others",
                           country == "Czech Republic" ~ "Others",
                           country == "The Netherlands" ~ "Others",
                           country == "Qatar" ~ "Others",
                           country == "Vietnam" ~ "Others",
                           country == "Brazil" ~ "Others",
                           country == "Spain" ~ "Others",
                           country == "Thailand" ~ "Others",
                           country == "Tunisia" ~ "Others",
                           country == "China" ~ "China",
                           country == "Belgium" ~ "Belgium",
                           country == "Pakistan" ~ "Pakistan",
                           country == "France" ~ "France",
                           country == "Australia" ~ "Australia",
                           country == "Germany" ~ "Germany",
                           country == "India" ~ "India",
                           country == "Canada" ~ "Canada",
                           country == "USA" ~ "USA",
                           country == "Reunion" ~ "Reunion"))
data.1
```

Plot graph
```{r}
plot <- 
  ggplot(data.1, aes(x = continent, fill= newCountry)) + 
  geom_bar(color = "#023047", width = 0.5) +
  xlab("Continent")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("China"="#780000", "Belgium"="#b7e4c7", "Pakistan"="#ff4d6d", 
                               "France"="#95d5b2", "Germany"="#74c69d", "Australia"="#219ebc",
                               "India"="#f28482", "Canada"="#e09f3e", "USA"="#ffba08",
                               "Reunion"="#06d6a0",
                               na.value = "grey"))+
  theme_bw() +
  facet_wrap(~ sample_type)
plot

ggsave("figure_3b.pdf", plot, width = 8, height = 6)
```


# Figure 4

## 4a - tree

Run rhierbaps
```{r}
snp.matrix <- load_fasta("figure_4/core.aln.fasta")

hb.results.2 <- hierBAPS(snp.matrix, max.depth = 2, n.pops = 20, n.extra.rounds = Inf, quiet = TRUE)
baps_cluster_df <- hb.results.2$partition.df
# write.csv(baps_cluster_df, "efm_baps_cluster.csv", row.names = FALSE)
```

Load tree file
```{r}
iqtree <- ape::read.tree("figure_4/core.aln.treefile")
iqtree

# Check tree tips and rhierbaps 
iqtree$tip.label
hb.results.2$partition.df$Isolate

hb.results.2$partition.df$Isolate %in% iqtree$tip.label
iqtree$tip.label %in% hb.results.2$partition.df$Isolate

all(iqtree$tip.label %in% hb.results.2$partition.df$Isolate)
iqtree$tip.label[!iqtree$tip.label %in% hb.results.2$partition.df$Isolate]

# Plot tree
tree.3 <- ggtree(iqtree)
tree.3 <- tree.3 %<+% hb.results.2$partition.df
tree.3 <- tree.3 + geom_tippoint(size=0.5, aes(color = factor(`level 1`)))
tree.3

# Load metadata
sample_data <- read.csv("/Users/jbbeh/Downloads/_VicLRE/figure_4/ggtree_metadata.csv")
View(sample_data)

# Check whether the Isolate IDs in both files align correctly
sample_data$ID %in% iqtree$tip.label
iqtree$tip.label %in% sample_data$ID

#Add root and scale bar
tree_mid <- midpoint(iqtree)
tree_mid
```

Make tree with rhierbaps levels
```{r}
tree.3 <- ggtree(tree_mid, branch.length = 0.5) %<+% hb.results.2$partition.df +
  geom_tippoint(size=0.5, aes(color = factor(`level 1`)))
tree.3

# Plot tree with metadata
tree.4 <- tree.3 %<+% sample_data +
  geom_rootedge(rootedge = 0.0005) +   # add root 
  geom_treescale(y=20)
tree.4

# Heatmap 1: MDU
mdu <- data.frame("MDU" = sample_data[,c("MDU")])
rownames(mdu) <- sample_data$ID

h1 <- 
  gheatmap(tree.4, mdu, width = 0.03, color=NULL, offset=0.05,
           colnames = FALSE) +
  scale_fill_manual(name = "continent",
                    values = c("#fb8500","#264653"),
                    na.value="#264653",  # set colour of NA values
                    breaks = c("Yes", "No"),
                    labels = c("Yes", "No")) +
  theme(legend.position = "right")
h1

# Heatmap 2: continent
length(unique(sample_data$continent))
unique(sample_data$continent) #count total number of continents

continent <- data.frame("continent" = sample_data[,c("continent")])
rownames(continent) <- sample_data$ID

h2 <- h1 + new_scale_fill()
h3 <- 
  gheatmap(h2, continent, width = 0.03, color=NULL, offset=0.07,
           colnames = FALSE) +
  scale_fill_manual(name = "continent",
                    values = c("#f2cc8f","#81b29a","#3d405b","#e07a5f","#fb6f92"),
                    na.value="grey",  # set colour of NA values
                    breaks = c("Asia", "Oceania", "South America", "North America", "Europe"),
                    labels = c("Asia", "Oceania", "South America", "North America", "Europe")) +
  theme(legend.position = "right")
h3

# Heatmap 3: sample type
length(unique(sample_data$sample_type))
unique(sample_data$sample_type)

sample_data$sample_type <- as.factor(sample_data$sample_type) 
sample_type <- data.frame("sample_type" = sample_data[,c("sample_type")])
rownames(sample_type) <- sample_data$ID

h4 <- h3 + new_scale_fill()
h5 <- 
  gheatmap(h4, sample_type, width = 0.03, color="white", offset=0.09,
           colnames = FALSE) +
  scale_fill_manual(name = "sample type",
                    values = c("#840032","#002642", "#E59500", "#840032"),
                    na.value="white",  # set colour of NA values
                    breaks = c("clinical/human", "animal", "environment","human"),
                    labels = c("clinical/human", "animal", "environment","human"))
h5

# Heatmap 4: ST
## Find top 10 prevalent STs
result <- sample_data %>%
  count(ST) %>%
  arrange(desc(n))
result

sample_data$ST <- as.factor(sample_data$ST) 
sample_data$ST[!(sample_data$ST %in% c("80", "117", "22","761", "132", "1421", "17", "32", "156", "324"))] <- 'others'
ST <- data.frame("ST" = sample_data[,c("ST")])
rownames(ST) <- sample_data$ID

h6 <- h5 + new_scale_fill()
h7 <- 
  gheatmap(h6, ST, width = 0.03, color="white", offset=0.12,
           colnames = FALSE) +
  scale_fill_manual(name = "sample type",
                    values = c("#840032","#002642", "#E59500", "#669bbc", "#588157",
                               "#ff99c8", "#83c5be", "#9f86c0", "#4361ee", "#f72585"),
                    na.value="white",  # set colour of NA values
                    breaks = c("80", "117", "22","761", "132", 
                               "1421", "17", "32", "156", "324"),
                    labels = c("80", "117", "22","761", "132", 
                               "1421", "17", "32", "156", "324"))
h7

# Heatmap 5: CC
sample_data$CC <- as.factor(sample_data$CC) 
CC <- data.frame("CC" = sample_data[,c("CC")])
rownames(CC) <- sample_data$ID
unique(sample_data$CC)

h8 <- h7 + new_scale_fill()
h9 <- 
  gheatmap(h8, CC, width = 0.03, color=NULL, offset=0.13,
           colnames = FALSE) +
  scale_fill_manual(name = "CC",
                    values = c("red", "white", "white"),
                    na.value="white",  # set colour of NA values
                    breaks = c("17", "94", "-"),
                    labels = c("17", "94", "-"))
h9

# Heatmap 6: LR gene cfr
sample_data$cfr <- as.factor(sample_data$cfr) 
cfr <- data.frame("cfr" = sample_data[,c("cfr")])
rownames(cfr) <- sample_data$ID

h10 <- h9 + new_scale_fill()
h11 <- 
  gheatmap(h10, cfr, width = 0.03, color="white", offset=0.15,
           colnames = FALSE) +
  scale_fill_manual(name = "cfr",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h11

# Heatmap 7: LR gene cfr(B)
sample_data$cfrB <- as.factor(sample_data$cfrB) 
cfrB <- data.frame("cfrB" = sample_data[,c("cfrB")])
rownames(cfrB) <- sample_data$ID

h12 <- h11 + new_scale_fill()
h13 <- 
  gheatmap(h12, cfrB, width = 0.03, color="white", offset=0.17,
           colnames = FALSE) +
  scale_fill_manual(name = "cfrB",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h13

# Heatmap 8: LR gene cfr(D)
sample_data$cfrD <- as.factor(sample_data$cfrD) 
cfrD <- data.frame("cfrD" = sample_data[,c("cfrD")])
rownames(cfrD) <- sample_data$ID

h14 <- h13 + new_scale_fill()
h15 <- 
  gheatmap(h14, cfrD, width = 0.03, color="white", offset=0.19,
           colnames = FALSE) +
  scale_fill_manual(name = "cfrD",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h15

# Heatmap 9: LR gene optrA
sample_data$optrA <- as.factor(sample_data$optrA) 
optrA <- data.frame("optrA" = sample_data[,c("optrA")])
rownames(optrA) <- sample_data$ID

h16 <- h15 + new_scale_fill()
h17 <- 
  gheatmap(h16, optrA, width = 0.03, color="white", offset=0.21,
           colnames = FALSE) +
  scale_fill_manual(name = "optrA",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h17

# Heatmap 10: LR gene poxtA
sample_data$poxtA <- as.factor(sample_data$poxtA) 
poxtA <- data.frame("poxtA" = sample_data[,c("poxtA")])
rownames(poxtA) <- sample_data$ID

h18 <- h17 + new_scale_fill()
h19 <- 
  gheatmap(h18, poxtA, width = 0.03, color="white", offset=0.23,
           colnames = FALSE) +
  scale_fill_manual(name = "poxtA",
                    values = c("black"),
                    na.value="white",  # set colour of NA values
                    breaks = c("yes"),
                    labels = c("yes"))
h19

# Heatmap 11: van type
sample_data$van_operon <- as.factor(sample_data$van_operon) 
van_operon <- data.frame("van_operon" = sample_data[,c("van_operon")])
rownames(van_operon) <- sample_data$ID

h20 <- h19 + new_scale_fill()
h21 <- 
  gheatmap(h20, van_operon, width = 0.03, color="white", offset=0.25,
           colnames = FALSE) +
  scale_fill_manual(name = "poxtA",
                    values = c("#1B98E0", "#6E2594", "white", "white"),
                    na.value="white",  # set colour of NA values
                    breaks = c("vanA", "vanB", "vanD"),
                    labels = c("vanA", "vanB", "vanD"))+
  theme(legend.position = "NULL")
h21

ggsave("figure_4a.pdf", h21, width=8, height=6)
```


## 4b - barplots

Load data
```{r}
data <- read.csv("figure_4/efs-metadata.csv")
data
```

Count prevalent countries
```{r}
count_country <- 
  data %>% dplyr::count(country) %>%
  arrange(desc(n))
count_country
```

Then select only the top 10 countries and change others into "Others"
```{r}
data.1 <-
  data %>% 
  mutate(newCountry = case_when(
                           country == "Spain" ~ "Others",
                           country == "United Kingdom" ~ "Others",
                           country == "Colombia" ~ "Others",
                           country == "Malaysia" ~ "Others",
                           country == "Tunisia" ~ "Others",
                           country == "Brazil" ~ "Others",
                           country == "Switzerland" ~ "Others",
                           country == "Denmark" ~ "Others",
                           country == "Australia" ~ "Others",
                           country == "Lebanon" ~ "Others",
                           country == "Bangladesh" ~ "Others",
                           country == "Pakistan" ~ "Others",
                           country == "Russia" ~ "Others",
                           country == "South Africa" ~ "Others",
                           country == "China" ~ "China",
                           country == "Belgium" ~ "Belgium",
                           country == "USA" ~ "USA",
                           country == "France" ~ "France",
                           country == "South Korea" ~ "South Korea",
                           country == "Germany" ~ "Germany",
                           country == "Portugal" ~ "Portugal",
                           country == "United Arab Emirates" ~ "United Arab Emirates",
                           country == "Canada" ~ "Canada",
                           country == "Italy" ~ "Italy",
                           country == "Japan" ~ "Japan", ))
data.1
```

Plot graph
```{r}
plot <- 
  ggplot(data.1, aes(x = continent, fill= newCountry)) + 
  geom_bar(color = "#023047", width = 0.5) +
  xlab("Continent")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("China"="#780000", "Belgium"="#b7e4c7", "USA"="#ffba08", 
                               "France"="#95d5b2", "South Korea"="#e5989b", "Germany"="#74c69d", 
                               "Portugal"="#52b788", "United Arab Emirates"="#ef233c", 
                               "Canada"="#e09f3e", "Italy"="#40916c", "Japan"="#ffcad4",
                               na.value = "grey"))+
  theme_bw() +
  facet_wrap(~ sample_type)
plot

ggsave("figure_4b.pdf", plot, width = 8, height = 6)
```


# Figure 5

## 5a - efs map

Import data
```{r}
data <- read.csv("figure_5/efs_map.csv")
data
```

Count the total number of isolates for each country
```{r}
country_counts <- data %>% dplyr::count(country)
typeof(country_counts)
country_counts <- data.frame(country_counts)
country_counts
```

Plot map
```{r}
colours <- c("#fcbf49","#faa307","#f48c06","#dc2f02","#d00000") #own colour palette
colourPalette <- RColorBrewer::brewer.pal(5,'RdPu') # colour palette from RColorBrewer

limits <- range(0, 400)  # custom range
legend.limits <- seq(min(limits), max(limits), len = 6)  # splitting to make continuous variable

# Make plot
pdf("plot_v1.pdf", width=80, height=60)
sPDF <- joinCountryData2Map(country_counts, joinCode='NAME', nameJoinColumn='country')
map.efs <- 
  mapCountryData(sPDF, nameColumnToPlot='n', #variable to be plotted
                 catMethod= legend.limits, #use custom scale
                 numCats = numCats,
                 mapTitle = "Global distribution of LR genes in Enterococcus",
                 missingCountryCol='dark grey', 
                 colourPalette=colours,
                 addLegend = TRUE)
do.call(addMapLegend, c(map.efs, horizontal=TRUE, legendWidth=0.5))
dev.off()
```

## 5b - efm map

Import data
```{r}
data <- read.csv("figure_5/efm_map.csv")
data
```

Count the total number of isolates for each country
```{r}
country_counts <- data %>% count(country)
typeof(country_counts)
country_counts <- data.frame(country_counts)
View(country_counts)
```

Plot map
```{r}
colours <- c("#fcbf49","#faa307","#f48c06","#dc2f02","#d00000") #own colour palette
colourPalette <- RColorBrewer::brewer.pal(5,'RdPu') # colour palette from RColorBrewer

limits <- range(0, 100)  # custom range
legend.limits <- seq(min(limits), max(limits), len = 6)  # splitting to make continuous variable

# Make plot
pdf("plot_v1.pdf", width=80, height=60)
sPDF <- joinCountryData2Map(country_counts, joinCode='NAME', nameJoinColumn='country')
map.efm <- 
  mapCountryData(sPDF, nameColumnToPlot='n', #variable to be plotted
                 catMethod= legend.limits, #use custom scale
                 numCats = numCats,
                 mapTitle = "Global distribution of LR genes in Enterococcus",
                 missingCountryCol='dark grey', 
                 colourPalette=colours,
                 addLegend = TRUE)
do.call(addMapLegend, c(map.efm, horizontal=TRUE, legendWidth=0.5))
dev.off()
```


# Figure S1

Figure S1 was created manually with Powerpoint.


# Figure S2

```{r}
# Load data
df <- read.csv("figure_s2/mic_input.csv")
View(df)

df$LZD_MIC <- as.factor(df$LZD_MIC)
df$G2576T_copyNumber <- as.factor(df$G2576T_copyNumber)

# Make plot
plot <- 
  ggplot(df, aes(x = G2576T_copyNumber, fill= LZD_MIC)) + 
  geom_bar(color = "#023047", width = 0.5) +
  xlab("ST")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  guides(fill=guide_legend(title="Linezolid MIC"))+
  scale_fill_manual(values = c("four"="#c1121f", "three"="#f77f00", "two"="#fcbf49", 
                               na.value = "white"))+
  theme_bw()
plot

# Save plot
ggsave("figure_s2.pdf", plot, width = 4, height = 6)
```


# Figure S3

Load data
```{r}
data_efm <- read.csv("/Users/jbbeh/Downloads/_VicLRE/figure_s3/efm-mic-dist.csv")
data_efs <- read.csv("/Users/jbbeh/Downloads/_VicLRE/figure_s3/efs-mic-dist.csv")
```

Group data into summary 
```{r}
data.1 <- 
  data_efm %>% 
  group_by(Year, LZD_susc) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  ungroup()
data.1

data.2 <- 
  data_efs %>% 
  group_by(Year, LZD_susc) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  ungroup()
data.2

```

Plot dividing MIC into six categories - <=0.5, 1, 2, 4, 8 and >8
```{r}
# Group data into summaries
data_efm$LZD_MIC <- as.factor(data_efm$LZD_MIC)
data.1 <- 
  data_efm %>% 
  group_by(Year, LZD_MIC) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  ungroup()
data.1

data_efs$LZD_MIC <- as.factor(data_efs$LZD_MIC)
data.2 <- 
  data_efs %>% 
  group_by(Year, LZD_MIC) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  ungroup()
data.2

# E. faecium plot
desired_order <- c(">8", "8", "4", "2", "1", "<=0.5", "NA")
data.1$LZD_MIC <- factor(data.1$LZD_MIC, levels = desired_order)

plot.1 <- 
  data.1 %>% 
  ggplot(aes(x=Year, y=n, fill=LZD_MIC)) + 
  geom_col() + 
  scale_fill_manual(values=c("#9d0208", "#d90429", "#ffb703",  "#90a955", "#4f772d", "#386641", "grey"))+
  theme_bw()+
  labs(title=NULL, y="Number of isolates", x="Year") +
  theme(legend.text = element_text(size = 7)) 
plot.1

# E. faecalis plot
desired_order <- c(">8", "8", "4", "2", "1", "<=0.5", "NA")
data.2$LZD_MIC <- factor(data.2$LZD_MIC, levels = desired_order)

plot.2 <- 
  data.2 %>% 
  ggplot(aes(x=Year, y=n, fill=LZD_MIC)) + 
  geom_col() + 
  scale_fill_manual(values=c("#9d0208", "#d90429", "#ffb703",  "#90a955", "#4f772d", "#386641", "grey"))+
  theme_bw()+
  labs(title=NULL, y="Number of isolates", x="Year") +
  theme(legend.text = element_text(size = 7)) 
plot.2

# Joining two plots together (top-bottom)
combined_plot <- wrap_plots(plot.2, plot.1, nrow = 2)  # arrange top-bottom
combined_plot

# Save plot
ggsave("figure_s3.pdf", combined_plot, width=6, height=4, dpi=300)
```


# Figure S6

## 6a

```{r}
# Set your desired file path
my_file_path <- "/Users/jbbeh/Library/CloudStorage/OneDrive-TheUniversityofMelbourne/PhD/ANALYSIS/r-plasmapR/Paper1_LRE/DMG2302727_plasmapR/DMG2302727.gbk"

# Assign the file path to the variable
fl <- my_file_path

# Create plot
fl |> 
  read_gb() |> 
  plot_plasmid(name = "DMG2302727-plasmid")
```

Access plasmid features 
```{r}
plasmid <- fl |> 
  read_gb()

dat <- plasmid |> 
  as.data.frame()
dat

# Make plot from data frame 
dat[dat$type == "CDS", ] |> 
  plot_plasmid(name = "my_plasmid")
```

Rename CDS based on Bakta GFF3 output
```{r}
#sample code
dat[dat == "DJGJNLLM_00006"] <- "repB" #rename 'DJGJNLLM_00006' as 'repB' gene
dat[dat == "DJGJNLLM_00009"] <- "relaxase"
dat[dat == "DJGJNLLM_00010"] <- "MobC"
dat[dat == "yxdl"] <- "ABC transporter ATP-binding protein yxdl"
dat[dat == "stp"] <- "fexB"
dat[dat == "DJGJNLLM_00031"] <- "Conjugal transfer protein"

View(dat)
```

Add new row of data/genes
```{r}
# Create a new row of data
new_row <- data.frame(index = 34, name = "ori", type = "CDS", start = 3487, end = 4337, direction = 1)
# Add the new row to the data frame using rbind()
dat <- rbind(dat, new_row)

new_row_2 <- data.frame(index = 35, name = "Type I toxin-antitoxin system", type = "CDS", start = 4131, end = 4217, direction = 1)
dat <- rbind(dat, new_row_2)

new_row_3 <- data.frame(index = 36, name = "IS6 family transposase", type = "CDS", start = 9394, end = 10080, direction = 1)
dat <- rbind(dat, new_row_3)

new_row_4 <- data.frame(index = 37, name = "IS1216 transposase", type = "CDS", start = 14105, end = 14791, direction = 1)
dat <- rbind(dat, new_row_4)

new_row_5 <- data.frame(index = 38, name = "IS1216 transposase", type = "CDS", start = 17426, end = 18112, direction = 1)
dat <- rbind(dat, new_row_5)

new_row_6 <- data.frame(index = 39, name = "IS1216 transposase", type = "CDS", start = 21023, end = 21709, direction = -1)
dat <- rbind(dat, new_row_6)

new_row_7 <- data.frame(index = 40, name = "IS1216 transposase", type = "CDS", start = 22970, end = 23656, direction = 1)
dat <- rbind(dat, new_row_7)

View(dat)
```

Remove all unnamed CDS
```{r}
filtered_dat <- dat[!grepl("^DJGJNLLM", dat$name), ]  # remove all gene names that start with DJGJNLLM
filtered_dat

plasmid <- 
  plot_plasmid(filtered_dat, name = "my_plasmid_2")
plasmid
```

Save plot
```{r}
ggsave(filename = "figure_s6a.pdf", plot = plasmid, width = 6, height = 4, dpi = 300)
```

## s6b

Create plot
```{r}
# Set your desired file path
my_file_path <- "figure_s6/DMG2302731.gbk"

# Assign the file path to the variable
fl <- my_file_path

#Create plot
fl |> 
  read_gb() |> 
  plot_plasmid(name = "DMG2302731-plasmid")
```

Access plasmid features 
```{r}
plasmid <- fl |> 
  read_gb()

dat <- plasmid |> 
  as.data.frame()

# Make plot from data frame 
dat[dat$type == "CDS", ] |> 
  plot_plasmid(name = "my_plasmid")
```

Rename CDS based on Bakta GFF3 output
```{r}
#sample code
dat[dat == "OEHAJAJD_00016"] <- "Conjugal transfer protein TraF" #rename 'OEHAJAJD_00016'
dat[dat == "OEHAJAJD_00020"] <- "Conjugal transfer protein" 
dat[dat == "OEHAJAJD_00029"] <- "IS1216 transposase" 
dat[dat == "OEHAJAJD_00031"] <- "transposase" 
dat[dat == "OEHAJAJD_00036"] <- "IS30 family transposase" 
dat[dat == "OEHAJAJD_00037"] <- "IS21-like element" 
dat[dat == "ble"] <- "bleO" 
dat[dat == "OEHAJAJD_00040"] <- "TnpA transposase" 
dat[dat == "OEHAJAJD_00045"] <- "IS1216 transposase" 
dat[dat == "OEHAJAJD_00049"] <- "IS1216 transposase" 
dat[dat == "OEHAJAJD_00056"] <- "IS1216 transposase" 
dat[dat == "OEHAJAJD_00063"] <- "RepR" 
dat[dat == "erm_1"] <- "erm(B)" 
dat[dat == "OEHAJAJD_00074"] <- "ISL3-like element" 

View(dat)
```

Add new row of data/genes
```{r}
# Create a new row of data
new_row <- data.frame(index = 80, name = "ori", type = "CDS", start = 4490, end = 4536, direction = 1)
# Add the new row to the data frame using rbind()
dat <- rbind(dat, new_row)

new_row_2 <- data.frame(index = 81, name = "mobA", type = "CDS", start = 5024, end = 7009, direction = 1)
dat <- rbind(dat, new_row_2)

new_row_3 <- data.frame(index = 82, name = "erm(C)", type = "CDS", start = 62757, end = 62840, direction = 1)
dat <- rbind(dat, new_row_3)

new_row_4 <- data.frame(index = 83, name = "erm(C)", type = "CDS", start = 70473, end = 70541, direction = 1)
dat <- rbind(dat, new_row_4)

new_row_5 <- data.frame(index = 84, name = "erm(B)", type = "CDS", start = 70666, end = 71403, direction = 1)
dat <- rbind(dat, new_row_5)

View(dat)
```

Remove unwanted rows
```{r}
dat[-3, ] 
```

Remove all unnamed CDS
```{r}
filtered_dat <- dat[!grepl("^OEHAJAJD", dat$name), ]  #remove all gene names that start with OEHAJAJD
filtered_dat

plasmid <- 
  plot_plasmid(filtered_dat, name = "my_plasmid_2")
plasmid
```

Save plot
```{r}
ggsave(filename = "figure_s6b.pdf", plot = plasmid, width = 6, height = 4, dpi = 300)
```

